---
output: 
  pdf_document:
    fig_crop: true
    keep_tex: true
    number_sections: true
    latex_engine: xelatex
    fig_caption: yes
title: "Efficient Frontier"
subtitle: "How should my portfolio look like?"
author: "Ana Karapetrovic-Suput"
date: "`r format(Sys.time(), '%d. %B %Y')`"
---
\newpage 
\tableofcontents 
\listoftables
\newpage
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r code, include=FALSE, message=FALSE, warning=FALSE}
library(tidyquant) # To download the data
library(plotly) # To create interactive charts
library(timetk) # To manipulate the data series
library(tidyr)
library(dplyr)

#Choose maximum 8 different stocks
tick = c("MMM", "AI.PA", "GIS", "SPGI", "UPS", "WMT", "GE", "PFE") #c("GIS", "SPGI", "WMT", "GE", "PFE")

#Get quantitative data in tibble format
efficient_frontier <- function(tick){
  obj <- NULL
  price_data <- tq_get(tick,
                       from = '2015-01-01',
                       to = '2021-12-31',
                       get = 'stock.prices') #Get stock price from yahoo finance
  
  #New variable "year" in price_data
  price_data_year <-price_data %>%
    mutate(year = substring(date, 1,4))
  
  #Calculate mean price pear year
  means_year <- price_data_year %>%
    group_by(symbol,year) %>%
    summarise_at(vars(adjusted), list(~mean(.)))#, na.rm=TRUE)))
  #Remove stocks with missing data
  stocks_na <-means_year%>%
    group_by(symbol)%>%
    summarise_at(vars(adjusted), list(~any(is.na(.))))
  means_year = means_year[means_year[["symbol"]] %in% stocks_na[["symbol"]][!stocks_na[["adjusted"]]],]
  
  ##Dismiss stocks stocks with greater value in the first year than in the last 
  means_year_min <- means_year[means_year[["year"]]==min(means_year[["year"]]),]
  means_year_max <- means_year[means_year[["year"]]==max(means_year[["year"]]),]
  selected_vars <- means_year_max[["symbol"]][means_year_max[["adjusted"]] > means_year_min[["adjusted"]]]
  
  if(length(selected_vars)<=1){
    obj = (selected_vars)
  }else{
    
    #Select stocks with greater value in the last than in the first year
    price_data_selected = price_data[price_data[["symbol"]] %in% selected_vars,]
    #Group price_data by symbol (=tick)
    log_ret_tidy <- price_data_selected %>%
      group_by(symbol) %>%
      #Calculates daily logarithmic periodic return
      tq_transmute(select = adjusted,
                   mutate_fun = periodReturn,
                   period = 'daily',
                   col_rename = 'ret',
                   type = 'log')
    #head(log_ret_tidy)
    
    #Stocks as colnames and date as rownames
    log_ret_xts <- log_ret_tidy %>%
      spread(symbol, value = ret) %>%
      tk_xts(date_var = date)
    #head(log_ret_xts)
    
    #Calculate daily mean
    mean_ret <- colMeans(log_ret_xts, na.rm = TRUE)
    #print(round(mean_ret, 5))
    
    #Calculate covariance matrix and anualize it
    cov_mat <- cov(log_ret_xts, use="complete.obs") * 252
    print(round(cov_mat,4))
    
    #Calculate weights
    n = ncol(cov_mat)
    wts = matrix(NA, ncol = n, nrow = 11^n)
    for (i in 1:n){
      wts[,i] <- rep(seq(0,1, 0.1), each = 11^(n-i), times = 11^(i-1))
    }
    wts <- wts[rowSums(wts)==1,]
    wts
    
    port_returns <- vector(length = dim(wts)[1])
    port_risk <- vector(length = dim(wts)[1])
    sharpe_ratio <- vector(length = dim(wts)[1])
    for (i in 1:dim(wts)[1]) {
      
      # Portfolio returns
      
      port_ret <- sum(wts[i,] * mean_ret, na.rm = TRUE)
      port_ret <- ((port_ret + 1)^252) - 1
      # Storing Portfolio Returns values
      port_returns[i] <- port_ret
      
      
      # Creating and storing portfolio risk
      port_sd <- sqrt(t(wts[i,]) %*% (cov_mat  %*% (wts[i,])))
      port_risk[i] <- port_sd
      
      # Creating and storing Portfolio Sharpe Ratios
      # Assuming 0% Risk free rate
      
      sr <- port_ret/port_sd
      sharpe_ratio[i] <- sr
      
    }
    
    # Storing the values in the table
    portfolio_values <- tibble(Return = port_returns,
                               Risk = port_risk,
                               SharpeRatio = sharpe_ratio)
    
    #Converting matrix to tibble
    all_wts <- tk_tbl(wts, preserve_index = FALSE)
    colnames(all_wts) <- colnames(log_ret_xts)
    
    # Combing all the values together
    portfolio_values <- tk_tbl(cbind(all_wts, portfolio_values), preserve_index = FALSE)
    
    #Minimum variance portfolio
    min_var <- portfolio_values[which.min(portfolio_values$Risk),]
    obj$min_var = min_var
    
    #Plot weights
    vars <- names(min_var)[names(min_var) %in% c("Return", "Risk", "SharpeRatio") == FALSE]
    obj$vars = vars
    p <- min_var %>%
      gather(vars, key = Asset, ##sollte hier alles nehmen, aber es nimmt nur GIS und PFE
             value = Weights) %>%
      mutate(Asset = as.factor(Asset)) %>%
      ggplot(aes(x = Asset, y = Weights, fill = Asset)) +
      geom_bar(stat = 'identity') +
      theme_minimal() +
      labs(x = 'Assets', y = 'Weights', title = "Minimum Variance Portfolio Weights") +
      scale_y_continuous(labels = scales::percent) 
    
    #ggplotly(p)
    obj$plot_minvar = p
    
    #tangency portfolio (the portfolio with highest sharpe ratio)
    max_sr <- portfolio_values[which.max(portfolio_values$SharpeRatio),]
    obj$max_sr = max_sr
    
    p <- max_sr %>%
      gather(vars, key = Asset,
             value = Weights) %>%
      mutate(Asset = as.factor(Asset)) %>%
      ggplot(aes(x = Asset, y = Weights, fill = Asset)) +
      geom_bar(stat = 'identity') +
      theme_minimal() +
      labs(x = 'Assets', y = 'Weights', title = "Tangency Portfolio Weights") +
      scale_y_continuous(labels = scales::percent) 
    
    #ggplotly(p)
    obj$plot_maxsr = p
    obj$portfolio_values = portfolio_values
    
    #Efficient frontier
    p <- portfolio_values %>%
      ggplot(aes(x = Risk, y = Return, color = SharpeRatio)) +
      geom_point() +
      theme_classic() +
      scale_y_continuous(labels = scales::percent) +
      scale_x_continuous(labels = scales::percent) +
      labs(x = 'Annualized Risk',
           y = 'Annualized Returns',
           title = "Portfolio Optimization and Efficient Frontier") +
      geom_point(aes(x = Risk, 
                     y = Return), data = min_var, color = 'orange') +
      geom_point(aes(x = Risk,
                     y = Return), data = max_sr, color = 'red') +
      annotate('text', x = max_sr$Risk*1.1, y = max_sr$Return*1.1, label = "Tangency Portfolio", color = "red") +
      annotate('text', x = min_var$Risk*1.2, y = min_var$Return*1.5, label = "Minimum Variance Portfolio", color = "orange")
      #
    
    #ggplotly(p)
    obj$plot_efficient_frontier = p
  }
  return(obj)
  
}

res = efficient_frontier(tick = tick)


```

# Efficient Frontier
If you are thinking about investing in stocks, you will come across the term "efficient frontier". The efficient frontier is the set of optimal portfolios that offer the highest expected return for a defined level of risk or the lowest risk for a given level of expected return. If you choose the portfolio with the lower risk or the higher expected return depends how risk-averse or risk-taking you are as an investor. 
In the example below I chose `r length(tick)` stocks for my portfolio. The following stocks were used at the beginning: `r paste0(tick)`. You can replace the stocks with any other you like. For more information visit https://finance.yahoo.com/. There you can also find the symbols of different stocks. 

In the picture below you can see the "Minimum Variance Portfolio" as well as the "Tangency Portfolio". The "Minimum Variance Portfolio" gives you a relative low return, but you have a low risk. If you choose the "Tangency Portfolio" you can earn more money, but its riskier. 

```{r plot efficient frontier, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", fig.height=4, dpi=300, results="hide"}

res$plot_efficient_frontier
```

\newpage
# Minimum Variance Portfolio
The details for the portfolio with the lowest risk are shown below. It is the portfolio with the lowest variance. 

```{r table Minimum Variance Portfolio, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", fig.height=4, dpi=300, results="asis"}
knitr::kable(res$min_var, caption = "Minimum Variance Portfolio - Weights, Return, Risk and Sharpe Ratio", format.args = list(decimal.mark = ",", big.mark = "."))
```

```{r plot Minimum Variance Portfolio, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", fig.height=4, dpi=300, results="hide"}
res$plot_minvar
```

\newpage
# Tangency Portfolio
The details for the portfolio with the highest sharpe ratio are shown below. The Sharpe ratio seeks to characterize how well the return of an asset compensates the investor for the risk taken. When comparing two assets, the one with a higher sharpe ratio appears to provide better return for the same risk, which is usually attractive to investors.

```{r table Tangency Portfolio, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", fig.height=4, dpi=300, results="asis"}
knitr::kable(res$max_sr, caption = "Tangency Portfolio - Weights, Return, Risk and Sharpe Ratio", format.args = list(decimal.mark = ",", big.mark = "."))
```


```{r plot Tangency Portfolio, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", fig.height=4, dpi=300, results="hide"}
res$plot_maxsr
```
